---
title: "QTL mapping for all traits in DO project" 
output: html_notebook
---

# Install libraries:

```{r}

library(qtl)
library(gplots)
library(ggplot2)
library(readxl)
library(RColorBrewer)
library(qtl2)
library(qtl2convert)

library(tidyverse, warn.conflicts = FALSE) # install.packages("tidyverse", dependencies = T)
library(stringr) # install.packages("stringr", dependencies = T)
library(qtl2) # install.packages("qtl2", dependencies = T)
library(ensimplR) # remotes::install_github("churchill-lab/ensimplR", dependencies = T)
library(ppcor) # install.packages("ppcor", dependencies = T)



```


# Load RDS files from mapping

```{r}

pheno <- read.csv("C:\\Users\\mkeller3\\Desktop\\Kahn FAHFA analysis Jan2026\\rz_20221206_DO_Batch_1to9_MPK_ComBat_Corrected_badmiceremoved.csv")


pheno=as.data.frame(pheno)
colnames(pheno)[1]="mouse"
rownames(pheno)=pheno$mouse

## Mapping files updated by Kelly
genoprobs=readRDS("W:\\General\\Projects2\\Holland500\\mapping_files\\Grcm39\\alleleprobs_grcm39_islet_clean.rds")
map=readRDS("W:\\General\\Projects2\\Holland500\\mapping_files\\Grcm39\\cleaned_DO500_pmap_grcm39.rds")
K <- readRDS("W:\\General\\Projects2\\Holland500\\mapping_files\\Grcm39\\k_loco_DO_islet_cleaned_grcm39.rds")
markers <- read_csv("W:\\General\\Projects2\\Holland500\\mapping_files\\Grcm39\\marker_annotations_DO500_grcm39.csv")



query_func = create_variant_query_func("W:\\General\\mapping_files_for_mark\\cc_variants_grcm39.sqlite")
query_genes = create_gene_query_func(dbfile = "W:\\General\\mapping_files_for_mark\\mouse_genes_grcm39.sqlite")



```



Load RDS files containing scans

```{r}

QTL_scan_add <- readRDS("C:\\Users\\mkeller3\\Desktop\\Kahn FAHFA analysis Jan2026\\QTL_FAHFA_DO_adipose_additive_Jan2026.rds")
QTL_scan_sex <- readRDS("C:\\Users\\mkeller3\\Desktop\\Kahn FAHFA analysis Jan2026\\QTL_FAHFA_DO_adipose_sexint_Jan2026.rds")

```



# Generate list of trait names for lipids

```{r}


variables <- names(pheno)[c(6:ncol(pheno))]


```


# Make sure covars are reclassifed as factors

```{r}

pheno$Diet <- as.factor(pheno$diet_days)
pheno$Wave <- as.factor(pheno$DOwave)
pheno$Sex <- as.factor(pheno$Sex)
pheno$Batch <- as.factor(pheno$Batch)


addcovar1 = model.matrix(~Sex+Diet+Wave+Batch, data = pheno)[,-1]

addcovar_interactive1 = t(t(model.matrix(~Sex, data = pheno)[,-1]))



```



Find peak locations from saved RDS file for all scans
```{r}

thr=7

addcovar1_peaks1 <- find_peaks(scan1_output = QTL_scan_add, map = map, threshold = thr, prob=0.95)

print(addcovar1_peaks1)



```


Modified Get allele effects function
```{r}

get_alleleEffects <- function(peaks, markers, geno, kinship, pheno, covar){
 
  peaks <- merge(peaks, markers[,c('marker','chr','pos')], by = c('chr','pos'), all.x = TRUE)
 
  peaks[,LETTERS[1:8]] <- 0
  for(i in 1:nrow(peaks)){
    gp <- geno[,peaks$chr[i]]
    gp[[1]] <- gp[[1]][,,peaks$marker[i], drop = FALSE]
   
    peaks[i, LETTERS[1:8]] <- scan1blup(genoprobs = gp,
                                       kinship   = K[[peaks$chr[i]]],
                                        pheno     = pheno[,peaks$lodcolumn[i], drop = FALSE],
                                        addcovar  = covar)[1, LETTERS[1:8]]
  }
 
  return(peaks)
} 


```


Combine allele effects to peaks for addtive scan
```{r}

addcovar1_peaks2 <- get_alleleEffects(peaks = addcovar1_peaks1, markers = markers, geno = genoprobs, kinship = K, pheno = pheno, covar = addcovar1)

write.csv(addcovar1_peaks2, file = "C:\\Users\\mkeller3\\Desktop\\Kahn FAHFA analysis Jan2026\\Additive_QTL_alleles_FAHFA_DO_adipose.csv")


```







Plot results from additive and interactive scans

```{r fig.height=5,fig.width=15}

trait <- "FL32_0_PAHPA_peak_4"

ylim=c(0,15)
chr = c("X")
xlim = c(5,20)

plot_scan1(x = QTL_scan_add[,trait], ylim=ylim, map = map, col="black", main = trait)
plot_scan1(x = QTL_scan_sex[,trait], ylim=ylim, map = map, col="darkred", add = TRUE)


plot_scan1(x = QTL_scan_add[,trait], chr = chr, xlim=xlim, ylim=ylim, map = map, col="black", main = trait)
plot_scan1(x = QTL_scan_sex[,trait], chr = chr, xlim=xlim, ylim=ylim, map = map, col="darkred", add = TRUE)


```


Plot results from additive and interactive scans

```{r fig.height=8,fig.width=5}

trait <- "FL34_1_POHSA_peak_6"

ylim=c(0,13)
chr = c("X")
xlim = c(5,30)

plot_scan1(x = QTL_scan_add[,trait], chr = chr, xlim=xlim, ylim=ylim, map = map, col="black", main = trait)
plot_scan1(x = QTL_scan_sex[,trait], chr = chr, xlim=xlim, ylim=ylim, map = map, col="darkred", add = TRUE)


```


Plot results from additive and interactive scans

```{r fig.height=5,fig.width=15}

trait1 <- "FL36_1_SAHOA_peak_7"
trait2 <- "FL36_1_SAHOA_peak_6"

ylim=c(0,22)
chr = c(11)
xlim = c(5,20)

plot_scan1(x = QTL_scan_add[,trait1], ylim=ylim, map = map, col="black", main = trait)
plot_scan1(x = QTL_scan_sex[,trait1], ylim=ylim, map = map, col="darkred", add = TRUE)

plot_scan1(x = QTL_scan_add[,trait2], ylim=ylim, map = map, col="black", main = trait)
plot_scan1(x = QTL_scan_sex[,trait2], ylim=ylim, map = map, col="darkred", add = TRUE)



```





Generate SNP association plot at specific peak
```{r fig.height=8,fig.width=15}

chr = "X"
start = 5
end = 11

assoc = scan1snps(genoprobs = genoprobs[,chr],
                  map = map,
                  pheno = pheno[,trait,drop = FALSE],
                  kinship = K,
                  addcovar = addcovar1,
                  intcovar = addcovar_interactive1, #sex
                  query_func = query_func,
                  chr = chr,
                  start = start,
                  end = end,
                  keep_all_snps = TRUE)

genes = query_genes(chr, start, end)
genes=genes[genes$stop-genes$start<1,]

plot_snpasso(assoc$lod,
             assoc$snpinfo,
             drop_hilit=1.5,
             main = trait,
             genes = genes,
             sdp_panel=TRUE,
             cex=1,
             color = "black",
             show_all_snps=TRUE)


```



# Generate a table containing SNPs at a QTL
```{r}


top <- top_snps(assoc$lod,
                assoc$snpinfo,
                lodcolumn = 1,
                chr = NULL,
                drop = 1.5,
                show_all_snps = TRUE)

write.csv(top, file = "C:/Users/mkeller3/Desktop/top_snps_FAHFA_QTLChr8.csv")



```



Create LOD difference profiles: sex_int - additive

```{r}

qtl_diff <- list()

qtl_peak <- as.data.frame(matrix(ncol = 7, nrow = 0)) 
colnames(qtl_peak)<- c("lodindex", "lodcolumn", "chr", "pos", "lod", "ci_lo", "ci_hi")

for (trait in variables) {
    
    # get peaks
    lod_diff <- QTL_scan_sex[, trait, drop=F] - QTL_scan_add[, trait, drop=F]
    peak_all <- find_peaks(lod_diff[, trait, drop=F], map, threshold = 5, prob=0.95)
    qtl_diff[[trait]]<-lod_diff
    
    if (nrow(peak_all) > 0 ) {
        qtl_peak <<- rbind(qtl_peak, peak_all)
    }
    
}

qtl_diff_all <- do.call("cbind", qtl_diff)

write.csv(qtl_peak, file = "C:/Users/mkeller3/Desktop/FAHFA_QTL_peak_LOD_difference_sex.csv")

saveRDS(qtl_diff_all, file = "C:/Users/mkeller3/Desktop/FAHFA_QTL_peak_LOD_difference_sex.rds")


```



Find peaks for all interactive scans

```{r}


add <- find_peaks(scan1_output = QTL_scan_add, map = map, threshold = 7, prob=0.95)
print(add)

sex <- find_peaks(scan1_output = QTL_scan_sex, map = map, threshold = 9, prob=0.95)
print(sex)


```





Make coeficient plots
```{r}
chr = 8
coef8 = scan1blup(genoprobs = genoprobs[,chr], pheno = pheno[,"FL36_1_SAHOA_peak_7", drop = FALSE], kinship = K[[chr]], addcovar = addcovar1)


```





Plot coeficients
```{r fig.height=6,fig.width=6}

plot_coefCC(x = coef8, ylim=c(-1.25,1.25), xlim=c(0,50), legend = "topright", map = map, main = "FL36_1_SAHOA_peak_7")


```





Combine allele effects to peaks
```{r}

addcovar1_peaks2 <- get_alleleEffects(peaks = addcovar1_peaks1, markers = markers, geno = genoprobs, kinship = K, pheno = pheno, covar = addcovar1)

write.csv(addcovar1_peaks1, file = "C:/Users/mkeller3/Desktop/Liver_lipid_QTL_sex.csv")


```

















